Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-11 ICPP-19/communications.tex
rearranging few sections

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with limited memory capacity, obtaining the optimal order of data transfers is an NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges that come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses and parallel resources, it is challenging to obtain a precise model based on prediction of computation and communication timings. Fourth, then number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tuned optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form Inria Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAAPI~\cite{kaapi} from Inria Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information is also provided along with the DAG, such as distance to exit (last) node as a priority or affinity of computation towards resources. The runtime is then responsible for managing scheduling of computations and communications, data transfers among different memories, computation-communication overlap, and load balance.

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities to computations, based on the task graph structure, and to use them at runtime to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance: tasks are added recursively based on certain sentinel constraints. For such applications, the runtime system sees a set of independent ready tasks and schedules them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations so as to maximize the overlap between computations and communications. This is the main topic of this paper. We prove that the order of communications on two memory nodes with the objective of minimizing the makespan is a NP-Complete problem if the memory of the target node is limited. Our proof is inspired from work by~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem with bounded capacity. The main difference between both approaches is that they consider all tasks have the same occupation on the second machine and the memory occupation starts when the processing finishes on the first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and we consider that memory is acquired before starting the data transfer on communication resource. We propose different runtime strategies in order to maximize the overlap of computations and communications. We evaluate our strategies on the context of a cluster of homogeneous nodes. However, our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper:

@@

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. We evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:exp}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem considered in this paper also can be viewed as a flow shop problem: the communication link can be seen as a processing resource, and each task needs to first be handled by the communication link and then by the computational resource. Communication and computation timings of a task can thus be considered as processing times on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}, in which a constraint is imposed on the number of tasks that can await execution on the second machine.

@@

	\todo[inline]{theoretical/performance work on molecular chemistry kernels}

@@

	Given a schedule, \scomm($i$) and \scomp($i$) represent the
	start times of task $i$ on communication and computation
	resources. A schedule is feasible if for every time $t$, the
	amount of memory required by all tasks such that $\scomm(i) \leq t
	\leq \scomp(i) + COMP_i$ is not more than the memory capacity
	$C$. 
	\todo[inline]{SKumar: We do not need this assumption throughout the paper. CommVolume=CommTime is not true for experimental section. We used this assumption only in NP-hard proof for simplicity.}
	For simplicity, we assume in all of this paper that the
	amount of memory required by a task is proportional to its
	communication time, and without loss of generality we consider
	that memory size of a task is equal to its communication time.

	We call a task $i$ compute intensive if $COMP_i \ge CM_i$, and
	communication intensive otherwise.

@@

		As \textit{Transfer time = Latency + (Data Volume)/Bandwidth}. For simplicity, we consider input transfer time is same as input data volume (i.e., \textit{Latency = 0} and \textit{Bandwidth = 1}). Otherwise, we can adjust $C$ such that at any point in a schedule at max one $K_i$ and three $A_i$ tasks can be active.

@@

		If the \threepart instance has a solution, $A$ can be partitioned into
		$m$ triplets $TR_i = \{a_{i_1}, a_{i_2}, a_{i_3}\}$ such that $\forall
		i, a_{i_1} + a_{i_2} + a_{i_3} = b$, then we can construct a feasible
		schedule $S$ without idle times by the pattern depicted in
		Figure~\ref{fig:firstSegment}. The communications of tasks in $TR_i$ take
		place during the computation of task $K_{i-1}$, and the computations
		of tasks in $TR_i$ take place during the communication of task
		$K_i$. Since the memory capacity is $C=b'+3$, all tasks from a triplet
		can fit in memory with a task $K_i$, and their duration are exactly
		equal to the communication time of $K_i$. This schedule is thus
		feasible, and has length exactly $L$.

@@

	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} presented in Section~\ref{sec:theoreticalProof} achieves an optimal makespan when there is no memory constraint. This optimal value indicates a lower bound on the makespan of the constrained case. We denote this value with $optimal\_ \allowbreak makespan\_  \allowbreak infinite\_  \allowbreak memory$ ($OMIM$). In the present Section, we propose different heuristics for the limited memory case, and we assess their efficiency with respect to this lower bound in Section~\ref{sec:exp}.

@@

	\subsection{Static Ordering}
	In this class of strategies, we compute the order of processing in advance based on criteria such as communication time and compute time. After computing the order, we follow the same order on computation and communication resources and make sure that the memory constraint is respected at each point in the schedule.

@@

	\begin{enumerate}[label=\roman*)]
		\item \textit{Order  of  optimal strategy infinite memory} ($OOSIM$): This heuristic uses the order given by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, but respects the memory constraint at each point in the schedule. Hence the makespan of this heuristic may be completely different from $OMIM$.

@@

	\subsection{Dynamic Selection}
	Dynamic strategies are based on the following principle: when the communication resource is idle, a task is chosen based on a ranking criterion which differ depending on the heuristic, among those which fit in memory. For example, if the selection criteria is to choose a highly compute intensive task, then we compute the ratio of compute time and communication time for all tasks, and we select the task which fit in the currently available memory and whose ratio is maximum (ties are broken arbitrarily). If no task satisfies the selection criteria thee we leave the resource idle at that point and proceed to the next event point. We also ensure that the order on both resources is the same, by ordering tasks on the computing resource accordingly.

@@

	%% We select a set of tasks which induce minimum idle time on compute resource and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		\item \textit{largest communication task respects memory restriction} ($LCMR$): a task with the largest communication time is chosen. 
		\item \textit{smallest communication task respects memory restriction} ($SCMR$): a task with the smallest communication time is chosen.
		\item \textit{maximum accelerated task respects memory restriction} ($MAMR$): a task with the maximum ratio of compute time to communication time is chosen.
	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

@@

		\centering
		\begin{tikzpicture}[yscale=0.6,xscale=0.3]
		\begin{scope}
		\schedule{24}{LCMR}{1,6,7,8,11,13,17,23}
		\taskB{0}
		\taskD[1]{1}
		\taskA{8}
		\taskC{13}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{26}{SCMR}{1,4,7,9,13,19,24,25}
		\taskB{0}
		\taskA[3]{1}
		\taskC{9}
		\taskD{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{25}{MAMR}{1,5,7,13,16,18,23,24}
		\taskB{0}
		\taskC[2]{1}
		\taskA{13}
		\taskD{18}
		\end{scope}
		\end{tikzpicture}
		\caption{ \label{fig:dynamicSelectionExample} Different dynamic heuristic schedules for a task set of Table~\ref{tab:dynamicSelectionExample} for a memory capacity of 6.}
		%%		 Optimal schedule length is 23.
	\end{figure}

@@

	In all strategies, the initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory. If not any task fits in the memory then we leave the resource idle and forward to the next event point.
%%	\todo[inline]{LED: What happens if no task fit in memory?}

	\begin{enumerate}[label=\roman*)]
		\item \textit{optimal order infinite memory largest communication task respects memory restriction} ($OOLCMR$): a task with the largest communication time is chosen from the set.
		\item \textit{optimal order infinite memory smallest communication task respects memory restriction} ($OOSCMR$): a task with the smallest communication time is chosen from the set.
		\item \textit{optimal order infinite memory maximum accelerated task respects memory restriction} ($OOMAMR$): a task  with the highest ratio of compute time to communication time is chosen from the set.
	\end{enumerate}

@@

	\begin{figure}[htb]
		\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
		\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
		\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
		\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
		\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
		\begin{tikzpicture}[yscale=0.6,xscale=0.2]
		\begin{scope}
		\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
		\taskB{0}
		\taskD[1]{2}
		\taskA{8}
		\taskE{12}
		\taskC{17}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{36}{OOSCMR}{2,5,8,11,14,18,26,34}
		\taskB{0}
		\taskE[3]{2}
		\taskA[1]{5}
		\taskD{9}
		\taskC{18}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
		\taskB{0}
		\taskD[1]{2}
		\taskE[1]{8}
		\taskA{12}
		\taskC{17}
		\end{scope}
		%%		\begin{scope}[yshift=+2.75cm]
		%%		\schedule{40}{OMIM}{2,8,16,24,29,33,38}
		%%		\taskB{0}
		%%		\taskC{8}
		%%		\taskD{24}
		%%		\taskA{29}
		%%		\taskE{33}
		%%		\end{scope}

		\end{tikzpicture}
		\caption{ \label{fig:staticOrderDynamicCorrectionsExample} Different static order dynamic corrections heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample} . The $OMIM$ order is $BCDAE$.}
	\end{figure}

@@

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of communications and computations. Recall that $COMP_i$ and $COMM_i$ represent computation and communication timings of task $i$, and that the memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent the start and end times of communication (resp. computation) for task $i$. The formulation also contains for each pair of tasks $i$ and $j$ \textit{i)} a boolean variable $a_{ij}$ to denote the order of $i$ and $j$ on the communication resource \textit{ii)} a boolean variables $b_{ij}$ to denote the order of $i$ and $j$ on the computation resource, and \textit{iii)} a boolean variables $c_{ij}$ to denote the order of $s_i$ and $e'_j$.

@@

	We use GLPK solver v4.65 to solve the above formulation. We also add the following constraints to help the solver: $\forall i, \forall j\ne i$, $a_{ij} + a_{ji}=1$, $b_{ij} + b_{ji} = 1$, $c_{ij} \le a_{ij}$, $c_{ij} \le b_{ij}$, and  $c_{ij} + c_{ji} \le 1$. The solver was unable to  solve this MILP at the scale of our interest in limited time. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$: at the boundary of two iterations we fix the event (communication or computation) of an unfinished task started before the boundary point and consider other events flexible. The subsets are formed in the order in which tasks are submitted, which is arbitrary. For a given size $k$, we represent the makespan calculated by this heuristic as $lp.k$.
	We compute various $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Hence, we do not include this heuristic for the comparison in Section~{\ref{sec:exp}. Performance comparison of different heuristics with MILP based heuristics is available in an extended version(cite).

@@

		\subsection{Favorable Situations for Heuristics}
		Based on the definition of proposed strategies and the optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, we present favorable scenarios for different heuristics in Table~\ref{tab:heuristicsAndFavorableScenarios}. This allows programmers to use appropriate strategies to maximize communication-computation overlap for their applications.

		\begin{table}[htb]
			\begin{tabular}{|c|p{6.5cm}|}
				\hline
				\textbf{Heuristic} & \textbf{\hspace{2cm}Favorable Situation} \\ \hline
				$OOSIM$ & Memory capacity is not a restriction (\textcolor{green}{Optimal}) \\ \hline
				$IOCMS$ & Memory capacity is not a restriction and tasks are compute intensive (\textcolor{green}{Optimal}) \\ \hline
				$DOCPS$ & Memory capacity is not a restriction and tasks are communication intensive (\textcolor{green}{Optimal}) \\ \hline
				$IOCCS$ & Moderate memory capacity and most tasks are highly compute intensive \\ \hline
				$DOCCS$ & Moderate memory capacity and most tasks are highly communication intensive \\ \hline
				$LCMR$ & Limited memory capacity and significant percentage of tasks with large communication timings are compute intensive\\ \hline
				$SCMR$ & Limited memory capacity and significant percentage of tasks with small communication timings are compute intensive\\ \hline
				$MAMR$ & Limited memory capacity and significant percentage of tasks of both types\\ \hline
				$OOLCMR$ & Moderate memory capacity and significant percentage of tasks are communication intensive\\ \hline
				$OOSCMR$ & Moderate memory capacity and significant percentage of tasks are compute intensive \\ \hline
				$OOMAMR$ & Moderate memory capacity and significant percentage of highly compute and communication intensive tasks \\ \hline
			\end{tabular}\caption{~\label{tab:heuristicsAndFavorableScenarios}Heuristics and their favorable scenarios.}
		\end{table}

>>>>

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		\todo[inline]{Initial submission deadline: April 15}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this paper, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with limited memory capacity, obtaining the optimal order of data transfers is a NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

@@

	With the advent of multicore, and the use of accelerators, it is notoriously cumbersome to exploit the full capability of a machine. Indeed, there are several challenges that come into picture. First, every architecture provides its own efficacy and interface. Therefore, a steep learning curve is required for programmers to take good utilization of all resources. Second, scheduling is a well known NP-Complete optimization problem, and hybrid and distributed resources make this problem harder (we refer ~\cite{webpagescheduling} for a survey on the complexity of scheduling problems and ~\cite{bleuse2015scheduling} for a recent survey in the case of hybrid nodes). Third, due to shared buses and parallel resources, it is challenging to obtain a precise model based on prediction of computation and communication times. Fourth, then number of architectures has increased drastically in recent years, therefore it is almost impossible to develop hand tuned optimized code for all these architectures. All these observations led to the development of different task based runtime systems. Among several runtimes, we may cite QUARK~\cite{YarKhan:2011:Quark:Manual} and PaRSEC~\cite{parsec} from from ICL, Univ. of Tennessee Knoxville (USA), StarPU~\cite{starpu} form Inria Bordeaux (France), Legion~\cite{legion12} from Stanford Univ. (USA), StarSs~\cite{ompss} from Barcelona Supercomputing Center (Spain), KAAPI~\cite{kaapi} from Inria Grenoble (France). All these runtime systems allow programmers to express their algorithms at the abstract level in the form of direct acyclic graphs (DAG), where vertices represent computations and edges represent dependencies among them. Sometimes some static information is also provided along with the DAG, such as distance to exit (last) node as a priority or affinity of computation towards resources. The runtime is then responsible for managing scheduling of computations and communications, data transfers among different memories, computation-communication overlap, and load balance.

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities to computations, based on the task graph structure, and to use them at runtime to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance: tasks are added recursively based on certain sentinel constraints. For such applications, the runtime system sees a set of independent tasks and schedules them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations so as to maximize the overlap between computations and communications. This is the main topic of this paper. We prove that the order of communications on two memory nodes with the objective of minimizing the makespan is a NP-Complete problem if the memory of the target node is limited. Our proof is inspired from work by~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem with bounded capacity. The main difference between both approaches is that they consider all tasks have the same occupation on the second machine and the memory occupation starts when the processing finishes on the first machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and we consider that memory is acquired before starting the data transfer on communication resource. We propose different runtime strategies in order to maximize the overlap of computations and communications. We evaluate our strategies on the context of a cluster of homogeneous nodes. However, our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper:

@@

	The outline of the paper is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. Sections~\ref{sec:expSetting} describes our experimental setup and we evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:expResults}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem considered in this paper also can be viewed as a flow shop problem: the communication link can be seen as a processing resource, and each task needs to first be handled by the communication link and then by the computational resource. Communication and computation times of a task can thus be considered as processing times on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}, in which a constraint is imposed on the number of tasks that can await execution on the second machine.

@@

%	\todo[inline]{theoretical/performance work on molecular chemistry kernels}

@@

	Given a schedule, \scomm($i$) and \scomp($i$) represent the
	start times of task $i$ on communication and computation
	resources. A schedule is feasible if for every time $t$, the
	amount of memory required by all tasks such that $\scomm(i) \leq t
	\leq \scomp(i) + COMP_i$ is not more than the memory capacity
	$C$. 
%%	\todo[inline]{SKumar: We do not need this assumption throughout the paper. CommVolume=CommTime is not true for experimental section. We used this assumption only in NP-hard proof for simplicity.}
%%	For simplicity, we assume in all of this paper that the
%%	amount of memory required by a task is proportional to its
%%	communication time, and without loss of generality we consider
%%	that memory size of a task is equal to its communication time.

 We assume through out the paper that communication time of a task is proportional to memory required by it, and follow \textit{Comm Time = Latency + (Comm Volume)/Bandwidth} model.

	We call a task $i$ compute intensive if $COMP_i \ge COMM_i$, and
	communication intensive otherwise.

@@

%%		As \textit{Transfer time = Latency + (Data Volume)/Bandwidth}. 
		For simplicity, we consider input transfer time is same as input data volume (i.e., \textit{Latency = 0} and \textit{Bandwidth = 1}). Otherwise, we can adjust $C$ such that at any point in a schedule at max one $K_i$ and three $A_i$ tasks can be active.

@@

		If the \threepart instance has a solution, $A$ can be partitioned into
		$m$ triplets $TR_i = \{a_{i_1}, a_{i_2}, a_{i_3}\}$ such that $\forall
		i, a_{i_1} + a_{i_2} + a_{i_3} = b$, then we can construct a feasible
		schedule $S$ without idle times by the pattern depicted in
		Figure~\ref{fig:firstSegment}. The communications of tasks in $TR_i$ take
		place during the computation of task $K_{i-1}$, and the computations
		of tasks in $TR_i$ take place during the communication of task
		$K_i$. Since the memory capacity is $C=b'+3$, all tasks from a triplet
		can fit in memory with a task $K_i$, and their durations are exactly
		equal to the communication time of $K_i$. This schedule is thus
		feasible, and has length exactly $L$.

@@

	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} presented in Section~\ref{sec:theoreticalProof} achieves an optimal makespan when there is no memory constraint. This optimal value indicates a lower bound on the makespan of the constrained case. We denote this value with \textit{optimal makespan infinite memory} ($OMIM$). In the present Section, we propose different heuristics for the limited memory case, and we assess their efficiency with respect to this lower bound in Section~\ref{sec:expResults}.

@@

	\subsection{Static Ordering}
	In this class of strategies, we compute the order of processing in advance based on criteria such as communication time and computation time. After computing the order, we follow the same order on computation and communication resources and make sure that the memory constraint is respected at each point in the schedule.

@@

	\begin{enumerate}[label=\roman*)]
		\item \textit{order  of  optimal strategy infinite memory} ($OOSIM$): This heuristic uses the order given by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, but respects the memory constraint at each point in the schedule. Hence the makespan of this heuristic may be completely different from $OMIM$.

@@

	\subsection{Dynamic Selection}
	Dynamic strategies are based on the following principle: when the communication resource is idle, a task is chosen based on a ranking criterion which differ depending on the heuristic, among those which fit in memory. For example, if the selection criteria is to choose a highly compute intensive task, then we compute the ratio of computation time and communication time for all tasks, and we select the task which induces minimum idle time on the computing resource, fit in the currently available memory and whose ratio is maximum (ties are broken arbitrarily). If no task fits in memory then we leave the resource idle at that point and proceed to the next event point. We also ensure that the order on both resources is the same, by ordering tasks on the computing resource accordingly.

@@

	%% We select a set of tasks which induce minimum idle time on compute resource and satisfy available memory requirement. Then a task is chosen for processing from this set based on the following heuristics.
	\begin{enumerate}[label=\roman*)]
		\item \textit{largest communication task respects memory restriction} ($LCMR$): a task with the largest communication time is chosen. 
		\item \textit{smallest communication task respects memory restriction} ($SCMR$): a task with the smallest communication time is chosen.
		\item \textit{maximum accelerated task respects memory restriction} ($MAMR$): a task with the maximum ratio of computation time to communication time is chosen.
	\end{enumerate}
	\begin{table}[htb]
		\begin{center}

@@

		\centering
		\begin{tikzpicture}[yscale=0.6,xscale=0.3]
		\begin{scope}
		\schedule{24}{LCMR}{1,6,7,8,11,13,17,23}
		\taskB{0}
		\taskD[1]{1}
		\taskA{8}
		\taskC{13}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
		\schedule{26}{SCMR}{1,4,7,9,13,19,24,25}
		\taskB{0}
		\taskA[3]{1}
		\taskC{9}
		\taskD{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{25}{MAMR}{1,5,7,13,16,18,23,24}
		\taskB{0}
		\taskC[2]{1}
		\taskA{13}
		\taskD{18}
		\end{scope}
		\end{tikzpicture}
		\caption{ \label{fig:dynamicSelectionExample} Different dynamic heuristic schedules for a task set of Table~\ref{tab:dynamicSelectionExample} with a memory capacity of 6.}
		%%		 Optimal schedule length is 23.
	\end{figure}

@@

	In all strategies, the initial order is obtained by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We define the following heuristics based on how we select a task from the set of tasks which fit in memory and induces minimum idle time on the computing resource. If no task fits in memory then we leave the resource idle and forward to the next event point.
	%%	\todo[inline]{LED: What happens if no task fit in memory?}

	\begin{enumerate}[label=\roman*)]
		\item \textit{optimal order infinite memory largest communication task respects memory restriction} ($OOLCMR$): a task with the largest communication time is chosen from the set.
		\item \textit{optimal order infinite memory smallest communication task respects memory restriction} ($OOSCMR$): a task with the smallest communication time is chosen from the set.
		\item \textit{optimal order infinite memory maximum accelerated task respects memory restriction} ($OOMAMR$): a task  with the highest ratio of computation time to communication time is chosen from the set.
	\end{enumerate}

@@

	\begin{figure}[htb]
		\newcommand{\taskA}[2][0]{\task[#1]{$A$}{#2}{4}{1}{cyan}}
		\newcommand{\taskB}[2][0]{\task[#1]{$B$}{#2}{2}{6}{cyan!50!black}}
		\newcommand{\taskC}[2][0]{\task[#1]{$C$}{#2}{8}{8}{blue!40!white}}
		\newcommand{\taskD}[2][0]{\task[#1]{$D$}{#2}{5}{4}{blue!70!white}}
		\newcommand{\taskE}[2][0]{\task[#1]{$E$}{#2}{3}{2}{blue}}
		\begin{tikzpicture}[yscale=0.6,xscale=0.2]
		\begin{scope}
		\schedule{35}{OOLCMR}{2,8,12,15,17,25,33}
		\taskB{0}
		\taskD[1]{2}
		\taskA{8}
		\taskE{12}
		\taskC{17}
		\end{scope}
		\begin{scope}[yshift=-2.75cm]
%%		\schedule{36}{OOSCMR}{2,5,8,11,14,18,26,34}
%%		\taskB{0}
%%		\taskE[3]{2}
%%		\taskA[1]{5}
%%		\taskD{9}
%%		\taskC{18}
%%		\end{scope}
		\schedule{36}{OOSCMR}{2,5,8,11,15,19,27,35}
		\taskB{0}
		\taskE[3]{2}
		\taskA[1]{5}
		\taskD{10}
		\taskC{19}
		\end{scope}
		\begin{scope}[yshift=-5.5cm]
		\schedule{36}{OOMAMR}{2,8,11,14,17,25,33}
		\taskB{0}
		\taskD[1]{2}
		\taskE[1]{8}
		\taskA{12}
		\taskC{17}
		\end{scope}
		%%		\begin{scope}[yshift=+2.75cm]
		%%		\schedule{40}{OMIM}{2,8,16,24,29,33,38}
		%%		\taskB{0}
		%%		\taskC{8}
		%%		\taskD{24}
		%%		\taskA{29}
		%%		\taskE{33}
		%%		\end{scope}

		\end{tikzpicture}
		\caption{ \label{fig:staticOrderDynamicCorrectionsExample} Different static order dynamic corrections heuristic schedules for a task set of Table~\ref{tab:staticOrderDynamicCorrectionsExample} with a memory capacity of 9. The $OMIM$ order is $BCDAE$.}
	\end{figure}

@@

We also consider two other static heuristics for evaluation. The first heuristic is based on an algorithm, proposed by Gilmore and Gomory, to obtain the minimal cost sequence for a set of jobs~\cite{Gilmore-Gomory:1964}. This is a classic algorithm for 2-machine no-wait flow shop. In this algorithm, each job has a start and end state and a cost is associated to change the state. In our context, this cost can be seen as non-overlap time of computation for two adjacent tasks. Here is the main idea of this algorithm. Initially, a partial sequence of jobs is represented by a graph such that their overlap is maximum. Subsequently edges are greedily added to this graph to connect two components while minimizing the total non-overlap cost. When the graph is  connected, an edge interchange mechanism is used to determine the sequence of jobs, which ensures that the sequence has minimal cost. More details can be found in the original paper~\cite{Gilmore-Gomory:1964} and our implementation is publicly available~\cite{gitworkrepo}. This algorithm does not take memory constraints into account and only provides the sequence of processing. We use this sequence with a memory capacity restriction just like for other static heuristcs, and we call this heuristic \textit{Gilmore-Gomory} ($GG$).

The second heuristic is based on the First-Fit algorithm for the bin packing problem. The idea of this heuristic is to identify groups of tasks which can fit in memory together, called \emph{bins}. In First-Fit, tasks are considered in an arbitrary order and added to the first bin in which they can fit. If no suitable bin is found then a new bin is created and this task is added to it. When all tasks have been assigned to bins, we consider the sequence made of all tasks from the first bin, then tasks for the second bin, and so on. We call this heuristic \textit{Bin Packing} ($BP$).

	\subsection{Solving Linear Program Iteratively}
	\label{subsec:linearprogrammingformulation}
	We use a mixed integer linear program to obtain the order of communications and computations. Recall that $COMP_i$ and $COMM_i$ represent computation and communication times of task $i$, and that the memory capacity of the target system is $C$. In the linear program formulation, $s_i$ and $e_i$ (resp. $s'_i$ and $e'_i$) represent the start and end times of communication (resp. computation) for task $i$. The formulation also contains for each pair of tasks $i$ and $j$ \textit{i)} a boolean variable $a_{ij}$ to denote the order of $i$ and $j$ on the communication resource \textit{ii)} a boolean variables $b_{ij}$ to denote the order of $i$ and $j$ on the computation resource, and \textit{iii)} a boolean variables $c_{ij}$ to denote the order of $s_i$ and $e'_j$.

@@

	We use GLPK solver v4.65 to solve the above formulation. We also add the following constraints to help the solver: $\forall i, \forall j\ne i$, $a_{ij} + a_{ji}=1$, $b_{ij} + b_{ji} = 1$, $c_{ij} \le a_{ij}$, $c_{ij} \le b_{ij}$, and  $c_{ij} + c_{ji} \le 1$. The solver was unable to  solve this MILP at the scale of our interest in limited time. Hence, we solve the linear program iteratively for a small subset of size $k=3,4,5$: at the boundary of two iterations we fix the event (communication or computation) of an unfinished task started before the boundary point and consider other events flexible. The subsets are formed in the order in which tasks are submitted, which is arbitrary. For a given size $k$, we represent the makespan calculated by this heuristic as $lp.k$.
	We compute various $lp.k$ values for different memory capacities and observe that most of the other heuristics perform better than this heuristic. Hence, we do not include this heuristic for the comparison in Section~{\ref{sec:expResults}. Performance comparison of different heuristics with MILP based heuristics is available in an extended version(cite).

@@

		\subsection{Favorable Situations for Heuristics}
		Based on the definition of proposed strategies and the optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, we present the scenarios which should be more favorable for each heuristic in Table~\ref{tab:heuristicsAndFavorableScenarios}. This allows programmers to use appropriate strategies to maximize communication-computation overlap for their applications. In this table, ``moderate memory capacity'' refers to the case where memory is constrained, but is close to the maximal memory requirement of the $OMIM$ schedule.
